# Multi-instrument variables {#multi_instrument_var}

## Date and time variables

As dates and times are used across instruments to capture delays between an event and another (e.g. between index cardiotoxicity and death), there are attributed a specific naming scheme:

-   They do not start by the instrument identifier

-   They have their own identifier `date` for dates and `ti` for times.

-   Time variables always have 3 components:

    -   The `ti` identifier

    -   The first event identifier

    -   The second event identifier

There is an exception to this naming rule: time between 2 cures of ICI (either previous line or current line), which has the long suffix `regimen_freq`.

For example, time between index cardiotoxicity and death is `ti_ic_death`. An event identifier can have a middle `_` if its very long (e.g. `cl1stctla4_other` for current line first dose of an unspecified anti-CTLA4, although this is quite a rare case.

> Time variables represent **delays** between two dates.

Of note, times can be expressed in days, weeks or years.

| var                    | instrument    | unit  | quick desc                                              |
|-------------|-------------|-------------|-------------------------------|
| `ti_tobaccoquit_ic`    | demo          | years | From tobacco quitting to IC                             |
| `ti_otx_ic`            | demo          | years | From organ transplant x to IC                           |
| `ti_plicistop_ic`      | demo          | days  | From previous line ICI stop (end of regimen) to IC      |
| `pl_drug_regimen_freq` | previous_line | weeks | From one cure to another, during an ICI therapy regimen |
| `ti_ic_1ststeroid`     | ic            | days  | From IC to first dose of steroids                       |
| `ti_cl1stnivo_ic`      | cl            | days  | From current line ICI-drug 1st infusion to IC           |
| `ti_ic_ihdischarge`    | ih            | days  | Length of hospital stay                                 |

## Time to cardiotoxicity onset from immune checkpoint inhibitor introduction

As the title suggests, the idea is to collect the time to onset between the **first dose** of immune checkpoint inhibitor received **during the current cycle** and the occurrence of cardiotoxicity.

```{r echo=FALSE}
library(ggplot2) 

data1 <- 
  data.frame(
  v1 = c(1, 2, 3, 4, 5),
  names = c(
    "start prev line",
    "stop prev line",
    "start current line",
    "(stop current line)",
    "cardiotoxicity index date"
  )
)

ggplot(data = data1, aes(x = v1)) +
  geom_text(aes(label = names),
             y = 1.2,
             angle = 30, 
            hjust = 0) +
  geom_segment(aes(xend = v1, y = 0, yend = 1), color = "grey") +
  geom_segment(
    x = min(data1$v1),
    xend = max(data1$v1) + 1,
    y = 0,
    yend = 0,
  color = "black",
  arrow = arrow(ends = "last", type = "closed")
) +
  geom_rect(
    aes(xmin = 3, xmax = 5, ymin = 0.3, ymax = .5), fill = "orange"
  ) +
  geom_text(x = 4, y = .4, label = "time to onset") +
  coord_cartesian(xlim = c(min(data1$v1), max(data1$v1 + 1)), ylim = c(-0.5, 2)) +
  theme_void()
```

This data is stored at the immune checkpoint inhibitor level, e.g., a patient who received nivolumab will have this data stored in a nivolumab specific variable. The structure for this data is

| Var name               | Description                                                                                        |
|---------------------|---------------------------------------------------|
| `ti_cl1stdrug_ic`      | It is a **delay**, in days. Drug name is abbreviated to a 4 digit name                             |
| `ti_cl1stdrug_ic__c`   | Same but computed from dates (when available).                                                     |
| `ti_cl1stdrug_ic__old` | It is an old variable which has now the \\\@ HIDDEN status, and is similar to `ti_cl1stdrug_ic__c` |

To capture this time to onset, you have to check every possibilities for a patient, e.g. you must check for all immune checkpoint inhibitor level time variables and choose which of these 3 variables you keep first, if ever more than one is available. This gives a quite heavy piece of code

```{r, eval=FALSE}
ti_cl1stici_ic = expr(pmax( # pmax here: maximum delay (in the case multiple ICI were prescribed sequentially)

    eval(tto_uni_3v(ti_cl1statez_ic__c, ti_cl1statez_ic, ti_cl1statez_ic__old)), 
    eval(tto_uni_3v(ti_cl1stavel_ic__c, ti_cl1stavel_ic, ti_cl1stavel_ic__old)), 
    eval(tto_uni_3v(ti_cl1stcemi_ic__c, ti_cl1stcemi_ic, ti_cl1stcemi_ic__old)), 
    eval(tto_uni_3v(ti_cl1stdurv_ic__c, ti_cl1stdurv_ic, ti_cl1stdurv_ic__old)), 
    eval(tto_uni_3v(ti_cl1stipil_ic__c, ti_cl1stipil_ic, ti_cl1stipil_ic__old)), 
    eval(tto_uni_3v(ti_cl1stnivo_ic__c, ti_cl1stnivo_ic, ti_cl1stnivo_ic__old)), 
    eval(tto_uni_3v(ti_cl1stpemb_ic__c, ti_cl1stpemb_ic, ti_cl1stpemb_ic__old)), 
    eval(tto_uni_3v(ti_cl1sttrem_ic__c, ti_cl1sttrem_ic, ti_cl1sttrem_ic__old)), 
    eval(tto_uni_3v(ti_cl1stpd1_other_ic__c, ti_cl1stpd1_other_ic, 
        ti_cl1stpd1_other_ic__old)), 
    eval(tto_uni_3v(ti_cl1stpdl1_other_ic__c, 
        ti_cl1stpdl1_other_ic, ti_cl1stpdl1_other_ic__old)), 
    eval(tto_uni_3v(ti_cl1stctla4_other_ic__c, ti_cl1stctla4_other_ic, 
        ti_cl1stctla4_other_ic__old)), 
    na.rm = TRUE)
  )
```

Where `eval(tto_univ_3v())` would be a prioritizer function among the 3 variables. Note that we might want to substract the `ti_icsy_ic` delay, if we're interested in the beginning of symptoms rather than index cardiotoxicity date.

```{r}
tto_uni_3v <- # A 3 variables time to onset unifier !
  function( # used to organize the three variables of time to cardiotox onset from drug initiation
    ti_icistart_event__c, # all quasi quoted names of columns from data. One is the calculated delay `ti_icistart_event__c`, one is manually entered when dates are not available `ti_icistart_event`, one is a old version of the variable `ti_icistart_event__old`
    ti_icistart_event,
    ti_icistart_event__old
  ){
    ti_icistart_event__c <- enexpr(ti_icistart_event__c)
    ti_icistart_event <- enexpr(ti_icistart_event)
    ti_icistart_event__old <- enexpr(ti_icistart_event__old)
    
    # underlying logic: 
    # first, prioritize calculated over manual vars with numvar_uni (intermediate object uni).
    # second, take the longest delay of uni and __old, if both are available.
    # third, if only one available, keep it (via na.rm = TRUE in pmax)
    
    # generate an expression to call numvar_uni
    
    uni <- numvar_uni(!!ti_icistart_event__c,
                      !!ti_icistart_event)
    
    ex <- quo(pmax(
      eval(uni),
      as.numeric(!!ti_icistart_event__old),
      na.rm = TRUE)
    )
    
    
    ex
  }

numvar_uni <- # Numeric variables unifier
  function( # used to organize data entered from 2 variables, currently its a prioritization
    var1, # all quasi quoted names of columns from data. usually, one is the calculated var, one is manually entered, var1 will be prioritize over var2
    var2
    # underlying data.frame data argument is omitted
  ){
    var1 <- enexpr(var1)
    var2 <- enexpr(var2)
    
    ex <- expr(case_when(
      !is.na(!!var1) ~ as.numeric(!!var1),
      !is.na(!!var2) ~ as.numeric(!!var2),
      TRUE ~ NA_real_
    ))
    
    ex
  }
```

## Death

-   `ih_ou_death` : During index hospitalization

-   `fu_death` : During follow-up (later than discharge)
