# Multi-instrument variables {#multi_instrument_var}

## Date and time variables

As dates and times are used across instruments to capture delays between an event and another (e.g. between index cardiotoxicity and death), there are attributed a specific naming scheme:

-   They do not start by the instrument identifier

-   They have their own identifier `date` for dates and `ti` for times.

-   Time variables always have 3 components:

    -   The `ti` identifier

    -   The first event identifier

    -   The second event identifier

There is an exception to this naming rule: time between 2 cures of ICI (either previous line or current line), which has the long suffix `regimen_freq`.

For example, time between index cardiotoxicity and death is `ti_ic_death`. An event identifier can have a middle `_` if its very long (e.g. `cl1stctla4_other` for current line first dose of an unspecified anti-CTLA4, although this is quite a rare case.

> Time variables represent **delays** between two dates.

Of note, times can be expressed in days, weeks or years.

+------------------------+---------------+-----------+---------------------------------------------------------+
| var                    | instrument    | unit      | quick desc                                              |
+========================+===============+===========+=========================================================+
| `ti_tobaccoquit_ic`    | demo          | years     | From tobacco quitting to IC                             |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `ti_otx_ic`            | demo          | years     | From organ transplant x to IC                           |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `ti_plicistop_ic`      | demo          | days      | From previous line ICI stop (end of regimen) to IC      |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `pl_drug_regimen_freq` | previous_line | weeks     | From one cure to another, during an ICI therapy regimen |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `ti_ic_1ststeroid`     | ic            | days      | From IC to first dose of steroids                       |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `ti_cl1stnivo_ic`      | cl            | days      | From current line ICI-drug 1st infusion to IC           |
+------------------------+---------------+-----------+---------------------------------------------------------+
| `ti_ic_ihdischarge`    | ih            | days      | Length of hospital stay                                 |
+------------------------+---------------+-----------+---------------------------------------------------------+
|                        |               |           |                                                         |
+------------------------+---------------+-----------+---------------------------------------------------------+
|                        |               |           |                                                         |
+------------------------+---------------+-----------+---------------------------------------------------------+
|                        |               |           |                                                         |
+------------------------+---------------+-----------+---------------------------------------------------------+
|                        |               |           |                                                         |
+------------------------+---------------+-----------+---------------------------------------------------------+

## Time to cardiotoxicity onset from immune checkpoint inhibitor introduction

As the title suggests, the idea is to collect the time to onset between the **first dose** of immune checkpoint inhibitor received **during the current cycle** and the occurrence of cardiotoxicity.

```{r echo=FALSE}
library(ggplot2) 

data1 <- 
  data.frame(
  v1 = c(1, 2, 3, 4, 5),
  names = c(
    "start prev line",
    "stop prev line",
    "start current line",
    "(stop current line)",
    "cardiotoxicity index date"
  )
)

ggplot(data = data1, aes(x = v1)) +
  geom_text(aes(label = names),
             y = 1.2,
             angle = 30, 
            hjust = 0) +
  geom_segment(aes(xend = v1, y = 0, yend = 1), color = "grey") +
  geom_segment(
    x = min(data1$v1),
    xend = max(data1$v1) + 1,
    y = 0,
    yend = 0,
  color = "black",
  arrow = arrow(ends = "last", type = "closed")
) +
  geom_rect(
    aes(xmin = 3, xmax = 5, ymin = 0.3, ymax = .5), fill = "orange"
  ) +
  geom_text(x = 4, y = .4, label = "time to onset") +
  coord_cartesian(xlim = c(min(data1$v1), max(data1$v1 + 1)), ylim = c(-0.5, 2)) +
  theme_void()
```

This data is stored at the immune checkpoint inhibitor level, e.g., a patient who received nivolumab will have this data stored in a nivolumab specific variable. The structure for this data is

+------------------------+-----------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------+
| Current                | Target                      | Description                                                                                                                                              |
+========================+=============================+==========================================================================================================================================================+
| `start_days_drug`      | `ti_icistart_ic_drug`       | It is a **delay**, in days. Drug name is abbreviated here in the current naming (e.g. nivo for nivolumab). Sometimes, the drug name is wrong (see below) |
|                        |                             |                                                                                                                                                          |
|                        |                             | Could be : time variable (ti\_) from (ici_start) to index cardiotoxicity (ic), for drug (drug complete name)                                             |
+------------------------+-----------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------+
| `start_days_drugx`     | `ti_icistart_ic_drug__calc` | Same but computed from dates (when available). Mind the "x" at the end of the name                                                                       |
|                        |                             |                                                                                                                                                          |
|                        |                             | Could be: same but with a standardized suffix indicating a calculation                                                                                   |
+------------------------+-----------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------+
| `first_doses_drug_mce` |                             | Drug name is complete here. It is an old variable which has now the \\\@ HIDDEN status, and is similar to start_days_drugx                               |
+------------------------+-----------------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------+

To capture this time to onset, you have to check every possibilities for a patient, e.g. you must check for all immune checkpoint inhibitor level time variables and choose which of these 3 variables you keep first, if ever more than one is available. This gives a quite heavy piece of code

```{r}
ti_icistart_ic = expr(pmax( # pmax here: maximum delay (in the case multiple ICI were prescribed sequentially)
   eval(tto_uni_3v(start_days_nivo,
                 start_days_nivox,
                 first_doses_nivolumab_mce)), # identical to start_days_nivox but it is @HIDDEN
   eval(tto_uni_3v(start_days_pem,
                 start_days_pemx,
                 first_doses_pembrolizumab_mce)), # @HIDDEN
   eval(tto_uni_3v(start_days_pem_2, # Other Anti-PD1 regimen
                 start_days_pem_2x,
                 first_doses_pembrolizumab_mce_2)),# @HIDDEN
   eval(tto_uni_3v(start_days_pem_3, # Cemiplimab 
                 start_days_pem_3x,
                 first_doses_pembrolizumab_mce_3)),
   eval(tto_uni_3v(start_days_atez,
                 start_days_atezx,
                 first_doses_atezolizumab_mce)),
   eval(tto_uni_3v(start_days_ave,
                 start_days_avex,
                 first_doses_avelumab_mce)),
   eval(tto_uni_3v(start_days_durv,
                 start_days_durvx,
                 first_doses_durvalumab_mce)),
   eval(tto_uni_3v(start_days_durv_3,# Ohter Anti-PDL1
                 start_days_durv_3x,
                 first_doses_durvalumab_mce_3)),
   eval(tto_uni_3v(start_days_ipi,
                 start_days_ipix,
                 first_doses_ipilimumab_mce)),
   eval(tto_uni_3v(start_days_trem,
                 start_days_tremx,
                 first_doses_tremelimumab_mce)),
   eval(tto_uni_3v(start_days_trem_2, # Other Anti-CTLA4
                 start_days_trem_2x,
                 first_doses_tremelimumab_mce_2)),
    na.rm = TRUE
  ) - # look at the minus!! Shall we keep it? 2022-09-16
    eval(numvar_uni(cardiotox_days_1,
                    cardiotox_days_1x))
    # Number of days before presentation that Myocarditis symptoms began
     
  )
```

Where `eval(tto_univ_3v())` would be a prioritizer function among the 3 variables. Note that we might want to substract the `cardiotox_days_1` delay, if we're interested in the beginning of symptoms rather than index date of referral.

## Death

-   `ih_ou_death` : During index hospitalization

-   `fu_death` : During follow-up (later than discharge)

## Electrocardiograms

-   `be` : Baseline (before immune checkpoint inhibitor current cycle 1st infusion)

-   `be2` : The same +++ it is a duplicate in a separate instrument (should be merged in the future). There are more features here.

-   `ie` : Index cardiotoxicity (1st EKG at the time of cardiotoxicity diagnosis)

-   `ie2` : The same ++ it is a duplicate in a separate instrument (should be merged in the future). There are more features here.

-   `ih_ekg` : Additional EKG(s) during index hospitalization

-   `fu_ekg` : Most recent EKG during follow-up

### Features

+---------------------------------+----------------------------------------------------------+
| Variable                        | Definition                                               |
+=================================+==========================================================+
| `exist`                         | Availability of said EKG                                 |
+---------------------------------+----------------------------------------------------------+
| `analysis`                      | All EKG diagnoses (rhythm, durations, repolarization...) |
+---------------------------------+----------------------------------------------------------+
| `cornell_voltage`               | Numeric features (only available for `be2` and `ie2`)    |
|                                 |                                                          |
| `sokolow_lyon_voltage`          |                                                          |
|                                 |                                                          |
| `heart_rate`                    |                                                          |
|                                 |                                                          |
| `pr_duration`                   |                                                          |
|                                 |                                                          |
| `qrs_duration`                  |                                                          |
|                                 |                                                          |
| `qt_duration (also qtcb, qtcf)` |                                                          |
+---------------------------------+----------------------------------------------------------+

Also, timings according to index cardiotoxicity, and upload variables.
